---
title: "Exercise 8"
author: "Laura Ye"
date: "September 20, 2017"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About the rnorm function

##What does the rnorm function do?

It has 3 inputs: n, mean, and ad. Change the values below to get a feel for how rnorm changes as these values change.

```{r, echo = FALSE}
library(shiny)
inputPanel(
  sliderInput("n", label = "n:",
              min = 10, max = 1000, value = 100, step = 10),
  sliderInput("mean", label = "mean:",
              min = -2, max = 2, value = 0, step = 0.1),
  sliderInput("sd", label = "sd:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

x0 <- 0
xL <- 20

renderPlot({
  y <- rnorm(input$n, input$mean, input$sd)
  plot(y, type="p", 
       main = paste0("rnorm Distribution (",input$n,",", input$mean, ",", input$sd,")"),
       ylab="Density",
       ylim = c(0,max(y)))
})
```

## Compare to a Normal Distribution

The normal distribution with the same mean and standard deviation would look like this: 

```{r, echo=FALSE}
x_min <- -5
x_max <- 5
y_min <- 0
y_max <- 1

x <- seq(x_min,x_max,length.out=100)

renderPlot({
  y <- dnorm(x, input$mean, input$sd)
  plot(x,y, type="l", 
       main = paste0("dnorm Distribution (", input$mean, ",", input$sd,")"),
       ylab="Density",
       xlim = c(x_min, x_max), 
       ylim = c(y_min, y_max))
})
```

## Histogram of rnorm

This is a plot of a histogram normalize to probability of n numbers generated by rnorm with the chose mean and standard deviation.

```{r, echo = FALSE}
renderPlot({
  y <- rnorm(input$n, input$mean, input$sd)
  hist(y, 
       main = paste0("histogram of rnorm Distribution (",input$n,",", input$mean, ",", input$sd,")"),
       ylab="Probability",
       freq=FALSE)
})
```


# Explanation of rnorm

As you can see in the graphs above, as n approaches a large number, the histogram with probabilities start to look more like a normal distribution with the same mean and deviation. This is because the rnorm function generates normal, but random variables within the given boundaries of mean and standard deviation. 

The mean dictates the *center* of the data. On the histogram, the mean will correspond to the x-value with the highest probability. 

The standard deviation dictates the *spread* of the data. With a high standard deviation, the spread of the data is large as well. With a low standard deviation, the spread of the data is small.