---
title: "Analysis of Eye of Argon"
author: "Laura Ye"
date: "January 15, 2018"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, encoding='ASCII')
```

## Data Source

Source of this analysis originates from a book by Jim Theis called "The Eye of Argon".

```{r link}
link_to_book <- 'https://ansible.uk/misc/eyeargon.html'
```

The book is stored as an html file and we will use ```RCurl``` to retrieve and ```XML``` to extract the text from this html file.

```{r extract}
library(RCurl)
library(XML)

#download html
html <- getURL(link_to_book, followlocation=TRUE)

#parse html
doc <- htmlParse(html, asText=TRUE)
plain_text <- xpathSApply(doc,
                          "//text()[not(ancestor::script)][not(ancestor::style)][not(ancestor::noscript)][not(ancestor::form)]", 
                          xmlValue)
plain_text <- iconv(plain_text, to = "ASCII//TRANSLIT")
```

## Cleaning the Data

To clean the data, we have to do the following:

- remove punctuation and invalid characters

- remove numbers

- remove empty strings

- convert to lowercase

- convert words to their stem (i.e. learned -> learn)

```{r clean}
#split strings into words and remove punctuation and numbers and invalid chars
words <- unlist(strsplit(plain_text, " |\r|\n|[-!?.,;:\\\"]+|[1234567890]+"))

#remove empty strings
words <- words [words != ""]

#convert to lowercase
words <- tolower(words)

library(SnowballC)
#convert to stem
word_stems <- wordStem(words, language="english")
```

Now that we've cleaned the data, we have a list of words appear in the story. Here is a preview of the list of words and word stems we will use to feed into the word cloud.

```{r preview_words}
words[1:25]
word_stems[1:25]
```

## Word Cloud

```{r word_cloud}
library(wordcloud)

wordcloud(words, max.words=100, random.order=FALSE)
wordcloud(word_stems, max.words=100, random.order=FALSE)
```

## Insights

We can see from the word cloud that "grignr" occurs the most frequently in this text, which is reasonable considering that the main character's name is Grignr. Other most commonly used words were upon, face, arms, head, face, hand, soldier, etc.  This is likely because the main character Grignr is very prone to describing people's features.