---
title: "Unsupervised Learning - Data Camp"
output: html_notebook
author: "Laura Ye"
---

# Welcome to the course!

Define unsupervised learning
Three types of machine learning
Execute one type of unsupervised learning

Unsupervised learning
- Data with no labels

Supervised Learning
- Making predictions based on labeled data
- regression or classification

Reinforcement learning
- Learns from feedback

Unsupervised LEarning
- find homogenous subgroups
- clustering to find subgroups
- segmenting customers or movies

Dimensionsity reduction
- patterns in features of data - dimensionality reduction
- visualization of high dimensional data
- preprocessing for supervised learning

challenges and benefits
- no single goal of analysis
- more unlabeled data than labeled

# Intro to k-means clustering

First of two clustering algorithms
Breaks observations into pre-defined number of clusters

kmeans(x, centers = 5, nstart=20)
centers = number of clusters
x = data
k means has a random component
- run multiple times to improve odds
nstart = number of times to repeat kmeans algorithm

## Exercises

k-means clustering

We have created some two-dimensional data and stored it in a variable called x in your workspace. The scatter plot on the right is a visual representation of the data.

In this exercise, your task is to create a k-means model of the x data using 3 clusters, then to look at the structure of the resulting model using the summary() function.

```{r}
# Create the k-means model: km.out
km.out <- kmeans(x, 3, nstart=20)

# Inspect the result
summary(km.out)
```

Results of kmeans()

The kmeans() function produces several outputs. In the video, we discussed one output of modeling, the cluster membership.

In this exercise, you will access the cluster component directly. This is useful anytime you need the cluster membership for each observation of the data used to build the clustering model. A future exercise will show an example of how this cluster membership might be used to help communicate the results of k-means modeling.

k-means models also have a print method to give a human friendly output of basic modeling results. This is available by using print() or simply typing the name of the model.

```{r}
# Print the cluster membership component of the model
print(km.out$cluster)

# Print the km.out object
print(km.out)
```

Visualizing and interpreting results of kmeans()

One of the more intuitive ways to interpret the results of k-means models is by plotting the data as a scatter plot and using color to label the samples' cluster membership. In this exercise, you will use the standard plot() function to accomplish this.

To create a scatter plot, you can pass data with two features (i.e. columns) to plot() with an extra argument col = km.out$cluster, which sets the color of each point in the scatter plot according to its cluster membership.

```{r}
# Scatter plot of x
plot(x, 
  col=km.out$cluster,
  main="k-means with 3 clusters",
  xlab="",
  ylab="")
```

# How kmeans() works

Visually understanding the algorithm
Model selection

Steps in algorithm
1. Randomly assign each point to one set of clusters.
2. Take the centers of the subgroups (average of all the positions) and each point is assigned to the closest center.
3. That was first iteration. This keeps going until no points are changing assignments.

As iterations increase, centers move further and further apart.

Model selection
- kmeans algorithm runs multiple times
- best outcome is based on total within cluster sum of squares
-- For each observation in each cluster, determine squared distance between the observation and the center
- use set.seed to guarantee reproducibility

scree plot
- run kmeans with different number of clusters, record total within ss and then plot
- elbow can be used to approximate the number of clusters

## Exercises

Handling random algorithms
In the video, you saw how kmeans() randomly initializes the centers of clusters. This random initialization can result in assigning observations to different cluster labels. Also, the random initialization can result in finding different local minima for the k-means algorithm. This exercise will demonstrate both results.

At the top of each plot, the measure of model quality???total within cluster sum of squares error???will be plotted. Look for the model(s) with the lowest error to find models with the better model results.

Because kmeans() initializes observations to random clusters, it is important to set the random number generator seed for reproducibility.

```{r}
# Set up 2 x 3 plotting grid
par(mfrow = c(2, 3))

# Set seed
set.seed(1)

for(i in 1:6) {
  # Run kmeans() on x with three clusters and one start
  km.out <- kmeans(x, centers=3, nstart=1)
  
  # Plot clusters
  plot(x, col = km.out$cluster, 
       main = km.out$tot.withinss, 
       xlab = "", ylab = "")
}
```

Selecting number of clusters
The k-means algorithm assumes the number of clusters as part of the input. If you know the number of clusters in advance (e.g. due to certain business constraints) this makes setting the number of clusters easy. However, as you saw in the video, if you do not know the number of clusters and need to determine it, you will need to run the algorithm multiple times, each time with a different number of clusters. From this, you can observe how a measure of model quality changes with the number of clusters.

In this exercise, you will run kmeans() multiple times to see how model quality changes as the number of clusters changes. Plots displaying this information help to determine the number of clusters and are often referred to as scree plots.

The ideal plot will have an elbow where the quality measure improves more slowly as the number of clusters increases. This indicates that the quality of the model is no longer improving substantially as the model complexity (i.e. number of clusters) increases. In other words, the elbow indicates the number of clusters inherent in the data.

```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# For 1 to 15 cluster centers
for (i in 1:15) {
  km.out <- kmeans(x, centers = i, nstart=20)
  # Save total within sum of squares to wss variable
  wss[i] <- km.out$tot.withinss
}

# Plot total within sum of squares vs. number of clusters
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Set k equal to the number of clusters corresponding to the elbow location
k <- 2
```

# Intro to Pokemon data

Gain some intuition about pokemon
hosted on kaggle

Challenges
- Selecting variables to cluster
- Scaling the data (usually to unite the uom)
- Determining the number of clusters
-- often no clean elbow in scree plot
- Visualize the results for interpretation

## Exercises

Practical matters: working with real data
Dealing with real data is often more challenging than dealing with synthetic data. Synthetic data helps with learning new concepts and techniques, but the next few exercises will deal with data that is closer to the type of real data you might find in your professional or academic pursuits.

The first challenge with the Pokemon data is that there is no pre-determined number of clusters. You will determine the appropriate number of clusters, keeping in mind that in real data the elbow in the scree plot might be less of a sharp elbow than in synthetic data. Use your judgement on making the determination of the number of clusters.

The second part of this exercise includes plotting the outcomes of the clustering on two dimensions, or features, of the data. These features were chosen somewhat arbitrarily for this exercise. Think about how you would use plotting and clustering to communicate interesting groups of Pokemon to other people.

An additional note: this exercise utilizes the iter.max argument to kmeans(). As you've seen, kmeans() is an iterative algorithm, repeating over and over until some stopping criterion is reached. The default number of iterations for kmeans() is 10, which is not enough for the algorithm to converge and reach its stopping criterion, so we'll set the number of iterations to 50 to overcome this issue. To see what happens when kmeans() does not converge, try running the example with a lower number of iterations (e.g. 3). This is another example of what might happen when you encounter real data and use real cases.

```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# Look over 1 to 15 possible clusters
for (i in 1:15) {
  # Fit the model: km.out
  km.out <- kmeans(pokemon, centers = i, nstart = 20, iter.max = 50)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Select number of clusters
k <- 2

# Build model with k clusters: km.out
km.out <- kmeans(x, centers = k, nstart = 20, iter.max = 50)

# View the resulting model
km.out

# Plot of Defense vs. Speed by cluster membership
plot(pokemon[, c("Defense", "Speed")],
     col = km.out$cluster,
     main = paste("k-means clustering of Pokemon with", k, "clusters"),
     xlab = "Defense", ylab = "Speed")
```

# Intro to hierarchical clustering

Used when number of clusters is not known
kmeans has to specify number of clusters
This course: bottom up clustering

Bottom up:
1. Assigns each group to its own cluster.
2. Find the closest two clusters and join into one.
3. Process continues iteratively by finding the next pair and combining.
4. Continues until there's only 1 cluster. Hierarchical clustering stops at this point.

hclus(d=dist_matrix)
- distance: euclidean (standard)

## Exercises

Hierarchical clustering with results
In this exercise, you will create your first hierarchical clustering model using the hclust() function.

We have created some data that has two dimensions and placed it in a variable called x. Your task is to create a hierarchical clustering model of x. Remember from the video that the first step to hierarchical clustering is determining the similarity between observations, which you will do with the dist() function.

You will look at the structure of the resulting model using the summary() function.

```{r}
# Create hierarchical clustering model: hclust.out
hclust.out <- hclust(dist(x))

# Inspect the result
summary(hclust.out)
```

# Selecting Number of Clusters

Interpreting results
- output is technical and opaque
- not that useful

Dendrogram
- Build a tree that represents the clustering process
- Distance between the two clusters is represented by height
-visualize: plot(hclust.out)

Choose number of clusters
- draw a horizontal line where you specify the height of the line, and you are drawing a line between the clusters
abline(h=6, col="red")

Tree cutting in R
- cut the tree to get cluster assignments
- specify height: cutree (hclust.out, h=6)
- specify num clusters: cutree(hclust.out, k=2)

## EXercises

Cutting the tree
Remember from the video that cutree() is the R function that cuts a hierarchical model. The h and k arguments to cutree() allow you to cut the tree based on a certain height h or a certain number of clusters k.

In this exercise, you will use cutree() to cut the hierarchical model you created earlier based on each of these two criteria.

```{r}
# Cut by height
cutree(hclust.out, h=7)

# Cut by number of clusters
cutree(hclust.out, k=3)
```

# Clustering linkage

How is distance between clusters determined? Rules?
Four Methods to measure which cluster should be linked
- Complete: pairwise similarity between all observations in both clusters - uses largest of similiarities
- SIngle: same as above but uses smallest of similiarities
- Average: same as above but uses average of similiarities
- Centroid: finds centroid of cluster 1 and centroid of cluster 2 and uses similarity between two centroids
Need to choose linkage type based on insights from the data
Rule of THumb: complete and average has more balanced trees and more commonly used.
Single: unbalanced trees
Centroid: Inversions (undesirable)

Linkage in R
specify method parameter
hclust.complete <- hclust(d, method = "complete"), etc

Scaling
- data on different scales can cause undesirable results in clustering
- solution is to scale so features have same mean and stderr
-- subtract mean from all observations
-- divide each feature by stderr
-- normalized features have a mean of zero and stderr of 1

Good idea to check the means and variability (stderr) between all the feateures

Check if scaling is necessary
colMeans(x) - mean of each column
apply(x, 2, sd) - find stderr of each column

scaled_x <- scale(x)
Check again with colMeans and apply

## Exercises

Linkage methods
In this exercise, you will produce hierarchical clustering models using different linkages and plot the dendrogram for each, observing the overall structure of the trees.

You'll be asked to interpret the results in the next exercise.

```{r}
# Cluster using complete linkage: hclust.complete
hclust.complete <- hclust(dist(x), method = "complete")

# Cluster using average linkage: hclust.average
hclust.average <- hclust(dist(x), method = "average")

# Cluster using single linkage: hclust.single
hclust.single <- hclust(dist(x), method = "single")

# Plot dendrogram of hclust.complete
plot(hclust.complete, main="Complete")

# Plot dendrogram of hclust.average
plot(hclust.average, main="Average")

# Plot dendrogram of hclust.single
plot(hclust.single, main="Single")
```

> Balanced trees are essential if you want an even number of observations assigned to each cluster. On the other hand, if you want to detect outliers, for example, an unbalanced tree is more desirable because pruning an unbalanced tree can result in most observations assigned to one cluster and only a few observations assigned to other clusters.

Practical matters: scaling
Recall from the video that clustering real data may require scaling the features if they have different distributions. So far in this chapter, you have been working with synthetic data that did not need scaling.

In this exercise, you will go back to working with "real" data, the pokemon dataset introduced in the first chapter. You will observe the distribution (mean and standard deviation) of each feature, scale the data accordingly, then produce a hierarchical clustering model using the complete linkage method.

```{r}
# View column means
colMeans(pokemon)

# View column standard deviations
apply(pokemon, 2, sd)

# Scale the data
pokemon.scaled <- scale(pokemon)

# Create hierarchical clustering model: hclust.pokemon
hclust.pokemon <- hclust(dist(pokemon.scaled), method = "complete")
```

> You first checked to see if the column means and standard deviations vary. Because they do, you scaled the data, converted the scaled data to a similarity matrix and passed it into the hclust() function

Comparing kmeans() and hclust()
Comparing k-means and hierarchical clustering, you'll see the two methods produce different cluster memberships. This is because the two algorithms make different assumptions about how the data is generated. In a more advanced course, we could choose to use one model over another based on the quality of the models' assumptions, but for now, it's enough to observe that they are different.

This exercise will have you compare results from the two models on the pokemon dataset to see how they differ.

```{r}
# Apply cutree() to hclust.pokemon: cut.pokemon
cut.pokemon <- cutree(hclust.pokemon, k = 3)

# Compare methods
table(km.pokemon$cluster, cut.pokemon)
```

hierarchical clustering model assigns most of the observations to cluster 1, while the k-means algorithm distributes the observations relatively evenly among all clusters. It's important to note that there's no consensus on which method produces better clusters. Judgement call.

# Intro to PCA

Unsupervised machine learning - dimensionality reduction
- find structure in features
- aid in visualization

Principal component analysis
3 goals
1. Find a linear combo of variables to create principal components - adds the features together
2. Maintain most variance in the data
3. PC are uncorrelated (orthogonal to each other)

PCA Intuition
2 dimensions (x and y)
Fina a lower dimensional state to describe the original data.
2D to 1D: 1 principal component.
Fit linear line to the data.
- First principal component of the data
Project values on the new line

Visualization of high dimensional data
- hard to visualize more than 3d
- can be easy to separate data using principal component
- first component retains 92% of the variance in the data

prcomp (x, scale, center)
x = data
scale = scale to 1 before PCA
center = center around zero (recommend leaving true)

summary(prcomp):
proportion of variance - variance explained by each PC
cumulative variance - variance as each component is added
standard deviation

## Exercises

PCA using prcomp()
In this exercise, you will create your first PCA model and observe the diagnostic results.

We have loaded the Pokemon data from earlier, which has four dimensions, and placed it in a variable called pokemon. Your task is to create a PCA model of the data, then to inspect the resulting model using the summary() function.

```{r}
# Perform scaled PCA: pr.out
pr.out <- prcomp(pokemon, scale=TRUE, center=TRUE)

# Inspect model output
summary(pr.out)
```

# Visualizing and interpreting PCA results

Biplot - ```biplot(pr.out)```
- PC2 vs PC1
- original features are correlated because dimensions are the same

scree plot
```{r}
pr.var <- pr.out$sdev^2 #square the stderr of each component (variance is defined as stderr^2)
pve <- pr.var / sum(pr.var) # variance proportion for each component
```
- show proportion of variance explained by each PC
- or cumulative variance
- when PC# = original dimension of the data, the proportion of variance explained is 1.0

## Exercises

Variance explained
The second common plot type for understanding PCA models is a scree plot. A scree plot shows the variance explained as the number of principal components increases. Sometimes the cumulative variance explained is plotted as well.

In this and the next exercise, you will prepare data from the pr.out model you created at the beginning of the chapter for use in a scree plot. Preparing the data for plotting is required because there is not a built-in function in R to create this type of plot.

```{r}
# Variability of each principal component: pr.var
pr.var <- pr.out$sdev^2

# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)
```

Visualize variance explained
Now you will create a scree plot showing the proportion of variance explained by each principal component, as well as the cumulative proportion of variance explained.

Recall from the video that these plots can help to determine the number of principal components to retain. One way to determine the number of principal components to retain is by looking for an elbow in the scree plot showing that as the number of principal components increases, the rate at which variance is explained decreases substantially. In the absence of a clear elbow, you can use the scree plot as a guide for setting a threshold.

```{r}
# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component",
     ylab = "Cumulative Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")
```

# Practical Issues with PCA

3 types of items to consider

1. Scaling data
2. Missing data (sometimes)
- not include or drop
- estimate or impute
3. Categorical data
- do not include
- encode categorical features as numbers

Scaling the data
- when data is in a different unit or order of magnitude, useful to scale or center
- Check biplot with scaling and without scaling
- with scaling, you find more even distribution of the loading vectors in pca
- without centering and scaling, two loading vectors might stick out because their order of magnitude is higher and hence variance is higher


## Exercises

Practical issues: scaling
You saw in the video that scaling your data before doing PCA changes the results of the PCA modeling. Here, you will perform PCA with and without scaling, then visualize the results using biplots.

Sometimes scaling is appropriate when the variances of the variables are substantially different. This is commonly the case when variables have different units of measurement, for example, degrees Fahrenheit (temperature) and miles (distance). Making the decision to use scaling is an important step in performing a principal component analysis.

```{r}
# Mean of each variable
colMeans(pokemon)

# Standard deviation of each variable
apply(pokemon, 2, sd)

# PCA model with scaling: pr.with.scaling
pr.with.scaling <- prcomp(pokemon, scale=TRUE)

# PCA model without scaling: pr.without.scaling
pr.without.scaling <- prcomp(pokemon, scale=FALSE)

# Create biplots of both for comparison
biplot(pr.with.scaling)
biplot(pr.without.scaling)
```

> The new Total column contains much more variation, on average, than the other four columns, so it has a disproportionate effect on the PCA model when scaling is not performed. After scaling the data, there's a much more even distribution of the loading vectors.

# Intro to case study

2 types of clustering
1 type of dimensionality reduction

Case study
- reinforce learning
- add steps not covered (preparing data)
- emphasize creativity

Example use case
Measuring nuclei of cells
Each record is ten features measured of each cell nuclei
Includes diagnosis (not used)

Steps
Download data and prepare
EDA
PCA + interpret
Two types of clustering
Understand and compare
Combine PCA and clustering

## Exercises

Preparing the data
Unlike prior chapters, where we prepared the data for you for unsupervised learning, the goal of this chapter is to step you through a more realistic and complete workflow.

Recall from the video that the first step is to download and prepare the data.

```{r}
url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1903/datasets/WisconsinCancer.csv"

# Download the data: wisc.df
wisc.df <- read.csv(url)

# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

# Create diagnosis vector
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```

Performing PCA
The next step in your analysis is to perform PCA on wisc.data.

You saw in the last chapter that it's important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data:

The input variables use different units of measurement.
The input variables have significantly different variances.

```{r}
# Check column means and standard deviations
colMeans(wisc.data)
apply(wisc.data, 2, sd)

# Execute PCA, scaling if appropriate: wisc.pr
wisc.pr <- prcomp(wisc.data, scale=TRUE)

# Look at summary of results
summary(wisc.pr)
```

Interpreting PCA results
Now you'll use some visualizations to better understand your PCA model. You were introduced to one of these visualizations, the biplot, in an earlier chapter.

You'll run into some common challenges with using biplots on real-world data containing a non-trivial number of observations and variables, then you'll look at some alternative visualizations. You are encouraged to experiment with additional visualizations before moving on to the next exercise.

```{r}
# Create a biplot of wisc.pr
biplot(wisc.pr)

# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[, c(1, 2)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC2")

# Repeat for components 1 and 3
plot(wisc.pr$x[, c(1, 3)], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")
```

>Because principal component 2 explains more variance in the original data than principal component 3, you can see that the first plot has a cleaner cut separating the two subgroups.

Variance explained
In this exercise, you will produce scree plots showing the proportion of variance explained as the number of principal components increases. The data from PCA must be prepared for these plots, as there is not a built-in function in R to create them directly from the PCA model.

As you look at these plots, ask yourself if there's an elbow in the amount of variance explained that might lead you to pick a natural number of principal components. If an obvious elbow does not exist, as is typical in real-world datasets, consider how else you might determine the number of principal components to retain based on the scree plot.

```{r}
# Set up 1 x 2 plotting grid
par(mfrow = c(1, 2))

# Calculate variability of each component
pr.var <- wisc.pr$sdev^2

# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")

# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")
```

> Before moving on, answer the following question: What is the minimum number of principal components needed to explain 80% of the variance in the data?

> 5

# PCA Review

Downloaded data and prepared for modeling
EDA
Performed detailed PCA

Next:
Complete hierarchical and kmeans clustering
Combine PCA and clustering
Constrast results of hierarchical and diagnosis
Compare Hier and Kmeans clustering
PCA as a pre-processing step for clustering

## Exercises

Hierarchical clustering of case data
The goal of this exercise is to do hierarchical clustering of the observations. Recall from Chapter 2 that this type of clustering does not assume in advance the number of natural groups that exist in the data.

As part of the preparation for hierarchical clustering, distance between all pairs of observations are computed. Furthermore, there are different ways to link clusters together, with single, complete, and average being the most common linkage methods.

```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)

# Calculate the (Euclidean) distances: data.dist
data.dist <- dist(data.scaled)

# Create a hierarchical clustering model: wisc.hclust
wisc.hclust <- hclust(data.dist, method="complete")
```

Selecting number of clusters
In this exercise, you will compare the outputs from your hierarchical clustering model to the actual diagnoses. Normally when performing unsupervised learning like this, a target variable isn't available. We do have it with this dataset, however, so it can be used to check the performance of the clustering model.

When performing supervised learning???that is, when you're trying to predict some target variable of interest and that target variable is available in the original data???using clustering to create new features may or may not improve the performance of the final model. This exercise will help you determine if, in this case, hierarchical clustering provides a promising new feature.

```{r}
# Cut tree so that it has 4 clusters: wisc.hclust.clusters
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)

# Compare cluster membership to actual diagnoses
table(diagnosis, wisc.hclust.clusters)
```

> Four clusters were picked after some exploration. Before moving on, you may want to explore how different numbers of clusters affect the ability of the hierarchical clustering to separate the different diagnoses.

k-means clustering and comparing results
As you now know, there are two main types of clustering: hierarchical and k-means.

In this exercise, you will create a k-means clustering model on the Wisconsin breast cancer data and compare the results to the actual diagnoses and the results of your hierarchical clustering model. Take some time to see how each clustering model performs in terms of separating the two diagnoses and how the clustering models compare to each other.

```{r}
# Create a k-means model on wisc.data: wisc.km
wisc.km <- kmeans(scale(wisc.data), 2, nstart=20)

# Compare k-means to actual diagnoses
table(wisc.km$cluster, diagnosis)

# Compare k-means to hierarchical clustering
table(wisc.hclust.clusters, wisc.km$cluster)
```

> Looking at the second table you generated, it looks like clusters 1, 2, and 4 from the hierarchical clustering model can be interpreted as the cluster 1 equivalent from the k-means algorithm, and cluster 3 can be interpreted as the cluster 2 equivalent.

Clustering on PCA results
In this final exercise, you will put together several steps you used earlier and, in doing so, you will experience some of the creativity that is typical in unsupervised learning.

Recall from earlier exercises that the PCA model required significantly fewer features to describe 80% and 95% of the variability of the data. In addition to normalizing data and potentially avoiding overfitting, PCA also uncorrelates the variables, sometimes improving the performance of other modeling techniques.

Let's see if PCA improves or degrades the performance of hierarchical clustering.


```{r}
# Create a hierarchical clustering model: wisc.pr.hclust
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method = "complete")

# Cut model into 4 clusters: wisc.pr.hclust.clusters
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)

# Compare to actual diagnoses
table(wisc.hclust.clusters, diagnosis)
table(wisc.pr.hclust.clusters, diagnosis)
table(wisc.km$cluster, diagnosis)

# Compare to k-means and hierarchical
table(wisc.pr.hclust.clusters, wisc.km$cluster)
```