---
title: "Unsupervised Learning - Data Camp"
output: html_notebook
author: "Laura Ye"
---

# Welcome to the course!

Define unsupervised learning
Three types of machine learning
Execute one type of unsupervised learning

Unsupervised learning
- Data with no labels

Supervised Learning
- Making predictions based on labeled data
- regression or classification

Reinforcement learning
- Learns from feedback

Unsupervised LEarning
- find homogenous subgroups
- clustering to find subgroups
- segmenting customers or movies

Dimensionsity reduction
- patterns in features of data - dimensionality reduction
- visualization of high dimensional data
- preprocessing for supervised learning

challenges and benefits
- no single goal of analysis
- more unlabeled data than labeled

# Intro to k-means clustering

First of two clustering algorithms
Breaks observations into pre-defined number of clusters

kmeans(x, centers = 5, nstart=20)
centers = number of clusters
x = data
k means has a random component
- run multiple times to improve odds
nstart = number of times to repeat kmeans algorithm

## Exercises

k-means clustering

We have created some two-dimensional data and stored it in a variable called x in your workspace. The scatter plot on the right is a visual representation of the data.

In this exercise, your task is to create a k-means model of the x data using 3 clusters, then to look at the structure of the resulting model using the summary() function.

```{r}
# Create the k-means model: km.out
km.out <- kmeans(x, 3, nstart=20)

# Inspect the result
summary(km.out)
```

Results of kmeans()

The kmeans() function produces several outputs. In the video, we discussed one output of modeling, the cluster membership.

In this exercise, you will access the cluster component directly. This is useful anytime you need the cluster membership for each observation of the data used to build the clustering model. A future exercise will show an example of how this cluster membership might be used to help communicate the results of k-means modeling.

k-means models also have a print method to give a human friendly output of basic modeling results. This is available by using print() or simply typing the name of the model.

```{r}
# Print the cluster membership component of the model
print(km.out$cluster)

# Print the km.out object
print(km.out)
```

Visualizing and interpreting results of kmeans()

One of the more intuitive ways to interpret the results of k-means models is by plotting the data as a scatter plot and using color to label the samples' cluster membership. In this exercise, you will use the standard plot() function to accomplish this.

To create a scatter plot, you can pass data with two features (i.e. columns) to plot() with an extra argument col = km.out$cluster, which sets the color of each point in the scatter plot according to its cluster membership.

```{r}
# Scatter plot of x
plot(x, 
  col=km.out$cluster,
  main="k-means with 3 clusters",
  xlab="",
  ylab="")
```

# How kmeans() works

Visually understanding the algorithm
Model selection

Steps in algorithm
1. Randomly assign each point to one set of clusters.
2. Take the centers of the subgroups (average of all the positions) and each point is assigned to the closest center.
3. That was first iteration. This keeps going until no points are changing assignments.

As iterations increase, centers move further and further apart.

Model selection
- kmeans algorithm runs multiple times
- best outcome is based on total within cluster sum of squares
-- For each observation in each cluster, determine squared distance between the observation and the center
- use set.seed to guarantee reproducibility

scree plot
- run kmeans with different number of clusters, record total within ss and then plot
- elbow can be used to approximate the number of clusters

## Exercises

Handling random algorithms
In the video, you saw how kmeans() randomly initializes the centers of clusters. This random initialization can result in assigning observations to different cluster labels. Also, the random initialization can result in finding different local minima for the k-means algorithm. This exercise will demonstrate both results.

At the top of each plot, the measure of model quality???total within cluster sum of squares error???will be plotted. Look for the model(s) with the lowest error to find models with the better model results.

Because kmeans() initializes observations to random clusters, it is important to set the random number generator seed for reproducibility.

```{r}
# Set up 2 x 3 plotting grid
par(mfrow = c(2, 3))

# Set seed
set.seed(1)

for(i in 1:6) {
  # Run kmeans() on x with three clusters and one start
  km.out <- kmeans(x, centers=3, nstart=1)
  
  # Plot clusters
  plot(x, col = km.out$cluster, 
       main = km.out$tot.withinss, 
       xlab = "", ylab = "")
}
```

Selecting number of clusters
The k-means algorithm assumes the number of clusters as part of the input. If you know the number of clusters in advance (e.g. due to certain business constraints) this makes setting the number of clusters easy. However, as you saw in the video, if you do not know the number of clusters and need to determine it, you will need to run the algorithm multiple times, each time with a different number of clusters. From this, you can observe how a measure of model quality changes with the number of clusters.

In this exercise, you will run kmeans() multiple times to see how model quality changes as the number of clusters changes. Plots displaying this information help to determine the number of clusters and are often referred to as scree plots.

The ideal plot will have an elbow where the quality measure improves more slowly as the number of clusters increases. This indicates that the quality of the model is no longer improving substantially as the model complexity (i.e. number of clusters) increases. In other words, the elbow indicates the number of clusters inherent in the data.

```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# For 1 to 15 cluster centers
for (i in 1:15) {
  km.out <- kmeans(x, centers = i, nstart=20)
  # Save total within sum of squares to wss variable
  wss[i] <- km.out$tot.withinss
}

# Plot total within sum of squares vs. number of clusters
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Set k equal to the number of clusters corresponding to the elbow location
k <- 2
```

# Intro to Pokemon data

Gain some intuition about pokemon
hosted on kaggle

Challenges
- Selecting variables to cluster
- Scaling the data (usually to unite the uom)
- Determining the number of clusters
-- often no clean elbow in scree plot
- Visualize the results for interpretation

## Exercises

Practical matters: working with real data
Dealing with real data is often more challenging than dealing with synthetic data. Synthetic data helps with learning new concepts and techniques, but the next few exercises will deal with data that is closer to the type of real data you might find in your professional or academic pursuits.

The first challenge with the Pokemon data is that there is no pre-determined number of clusters. You will determine the appropriate number of clusters, keeping in mind that in real data the elbow in the scree plot might be less of a sharp elbow than in synthetic data. Use your judgement on making the determination of the number of clusters.

The second part of this exercise includes plotting the outcomes of the clustering on two dimensions, or features, of the data. These features were chosen somewhat arbitrarily for this exercise. Think about how you would use plotting and clustering to communicate interesting groups of Pokemon to other people.

An additional note: this exercise utilizes the iter.max argument to kmeans(). As you've seen, kmeans() is an iterative algorithm, repeating over and over until some stopping criterion is reached. The default number of iterations for kmeans() is 10, which is not enough for the algorithm to converge and reach its stopping criterion, so we'll set the number of iterations to 50 to overcome this issue. To see what happens when kmeans() does not converge, try running the example with a lower number of iterations (e.g. 3). This is another example of what might happen when you encounter real data and use real cases.

```{r}
# Initialize total within sum of squares error: wss
wss <- 0

# Look over 1 to 15 possible clusters
for (i in 1:15) {
  # Fit the model: km.out
  km.out <- kmeans(pokemon, centers = i, nstart = 20, iter.max = 50)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
plot(1:15, wss, type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")

# Select number of clusters
k <- 2

# Build model with k clusters: km.out
km.out <- kmeans(x, centers = k, nstart = 20, iter.max = 50)

# View the resulting model
km.out

# Plot of Defense vs. Speed by cluster membership
plot(pokemon[, c("Defense", "Speed")],
     col = km.out$cluster,
     main = paste("k-means clustering of Pokemon with", k, "clusters"),
     xlab = "Defense", ylab = "Speed")
```

# Intro to hierarchical clustering

Used when number of clusters is not known
kmeans has to specify number of clusters
This course: bottom up clustering

Bottom up:
1. Assigns each group to its own cluster.
2. Find the closest two clusters and join into one.
3. Process continues iteratively by finding the next pair and combining.
4. Continues until there's only 1 cluster. Hierarchical clustering stops at this point.

hclus(d=dist_matrix)
- distance: euclidean (standard)

## Exercises

Hierarchical clustering with results
In this exercise, you will create your first hierarchical clustering model using the hclust() function.

We have created some data that has two dimensions and placed it in a variable called x. Your task is to create a hierarchical clustering model of x. Remember from the video that the first step to hierarchical clustering is determining the similarity between observations, which you will do with the dist() function.

You will look at the structure of the resulting model using the summary() function.

```{r}
# Create hierarchical clustering model: hclust.out
hclust.out <- hclust(dist(x))

# Inspect the result
summary(hclust.out)
```

# Selecting Number of Clusters

Interpreting results
- output is technical and opaque
- not that useful

Dendrogram
- Build a tree that represents the clustering process
- Distance between the two clusters is represented by height
-visualize: plot(hclust.out)

Choose number of clusters
- draw a horizontal line where you specify the height of the line, and you are drawing a line between the clusters
abline(h=6, col="red")

Tree cutting in R
- cut the tree to get cluster assignments
- specify height: cutree (hclust.out, h=6)
- specify num clusters: cutree(hclust.out, k=2)

## EXercises

Cutting the tree
Remember from the video that cutree() is the R function that cuts a hierarchical model. The h and k arguments to cutree() allow you to cut the tree based on a certain height h or a certain number of clusters k.

In this exercise, you will use cutree() to cut the hierarchical model you created earlier based on each of these two criteria.

```{r}
# Cut by height
cutree(hclust.out, h=7)

# Cut by number of clusters
cutree(hclust.out, k=3)
```

# Clustering linkage

How is distance between clusters determined? Rules?
Four Methods to measure which cluster should be linked
- Complete: pairwise similarity between all observations in both clusters - uses largest of similiarities
- SIngle: same as above but uses smallest of similiarities
- Average: same as above but uses average of similiarities
- Centroid: finds centroid of cluster 1 and centroid of cluster 2 and uses similarity between two centroids
Need to choose linkage type based on insights from the data
Rule of THumb: complete and average has more balanced trees and more commonly used.
Single: unbalanced trees
Centroid: Inversions (undesirable)

Linkage in R
specify method parameter
hclust.complete <- hclust(d, method = "complete"), etc

Scaling
- data on different scales can cause undesirable results in clustering
- solution is to scale so features have same mean and stderr
-- subtract mean from all observations
-- divide each feature by stderr
-- normalized features have a mean of zero and stderr of 1

Good idea to check the means and variability (stderr) between all the feateures

Check if scaling is necessary
colMeans(x) - mean of each column
apply(x, 2, sd) - find stderr of each column

scaled_x <- scale(x)
Check again with colMeans and apply

## Exercises

Linkage methods
In this exercise, you will produce hierarchical clustering models using different linkages and plot the dendrogram for each, observing the overall structure of the trees.

You'll be asked to interpret the results in the next exercise.

```{r}
# Cluster using complete linkage: hclust.complete
hclust.complete <- hclust(dist(x), method = "complete")

# Cluster using average linkage: hclust.average
hclust.average <- hclust(dist(x), method = "average")

# Cluster using single linkage: hclust.single
hclust.single <- hclust(dist(x), method = "single")

# Plot dendrogram of hclust.complete
plot(hclust.complete, main="Complete")

# Plot dendrogram of hclust.average
plot(hclust.average, main="Average")

# Plot dendrogram of hclust.single
plot(hclust.single, main="Single")
```

> Balanced trees are essential if you want an even number of observations assigned to each cluster. On the other hand, if you want to detect outliers, for example, an unbalanced tree is more desirable because pruning an unbalanced tree can result in most observations assigned to one cluster and only a few observations assigned to other clusters.

Practical matters: scaling
Recall from the video that clustering real data may require scaling the features if they have different distributions. So far in this chapter, you have been working with synthetic data that did not need scaling.

In this exercise, you will go back to working with "real" data, the pokemon dataset introduced in the first chapter. You will observe the distribution (mean and standard deviation) of each feature, scale the data accordingly, then produce a hierarchical clustering model using the complete linkage method.

```{r}
# View column means
colMeans(pokemon)

# View column standard deviations
apply(pokemon, 2, sd)

# Scale the data
pokemon.scaled <- scale(pokemon)

# Create hierarchical clustering model: hclust.pokemon
hclust.pokemon <- hclust(dist(pokemon.scaled), method = "complete")
```

> You first checked to see if the column means and standard deviations vary. Because they do, you scaled the data, converted the scaled data to a similarity matrix and passed it into the hclust() function

Comparing kmeans() and hclust()
Comparing k-means and hierarchical clustering, you'll see the two methods produce different cluster memberships. This is because the two algorithms make different assumptions about how the data is generated. In a more advanced course, we could choose to use one model over another based on the quality of the models' assumptions, but for now, it's enough to observe that they are different.

This exercise will have you compare results from the two models on the pokemon dataset to see how they differ.

```{r}
# Apply cutree() to hclust.pokemon: cut.pokemon
cut.pokemon <- cutree(hclust.pokemon, k = 3)

# Compare methods
table(km.pokemon$cluster, cut.pokemon)
```

hierarchical clustering model assigns most of the observations to cluster 1, while the k-means algorithm distributes the observations relatively evenly among all clusters. It's important to note that there's no consensus on which method produces better clusters. Judgement call.


