---  
title: "Assignment 1"
author: "LAURA YE"  
date: "October 11, 2017"  
output: 
  html_document:  
    toc: yes  

---  
email address: laura.ye.lin@gmail.com
  
***
Instructions: answer the questions below,  Do not use the R interpreter or the Internet during the quiz  

Use only base R code in your answers , no use of packages unless specified.  No use of for (or any other type of) loops on any questions.

(additional instructions and/or comments)


# Problem 1  
  
Given a 10 element numeric vector v, give an R statement that adds the mean of v to every individual element of v and assigns the result to v1  

```{r, comment=NA}
v <- 1:10
v_avg <- mean(v)
v_avg
v1 <- v_avg + v
v1
```
  
# Problem 2  
  
Given a character vector vchars, write a single statement that prints the number of missing values in vchar  
  
```{r, comment=NA}
vchars <- c("hello", "goodbye", "nothing", NA, NA, "see ya")
sum(is.na(vchars))
```

# Problem 3

Given a numeric vector nvec, write a single R statement that assigns a zero to every negative element present in nvec 

```{r, comment=NA}
nvec <- rnorm(10, 0, 5)
nvec
nvec[nvec<0] <- 0
nvec

```

# Problem 4

Given three vectors a, b, and d, create a fourth vector consisting of the concatenation of d, b, and a in that order

```{r, comment=NA}
a <- 0:10
b <- 5:10
d <- 20:25
c(d,b,a)
```

# Problem 5

What is printed by the following R statement?

print(typeof(c(TRUE, 317, 412, "2007-03-01"))

```{r, comment=NA}
"character"
```

# Problem 6

A data frame df has the following columns:

$ scores <dbl> 79.51508, 67.84057, 74.89246, 75.71270, 74.18168, 75.73072, 77.76586, 76.790...

$ eval   <chr> "k", "p", "i", "d", "c", "z", "m", "j", "t", "u"

(A) Write a single R statement that deletes all the rows in df for which scores is less than the mean of scores

```{r, comment=NA}
scores <- c(79.51508, 67.84057, 74.89246, 75.71270, 74.18168, 75.73072, 77.76586, 76.790)
eval <- c("k", "p", "i", "d", "c", "z", "m", "j")

df <- data.frame(scores,eval)
df
df_subset <- subset(df, df$scores>mean(scores))
df_subset
```

(B) Add a new column names new_scores to df consisting of a normally distributed random vector with mean 100 and standard deviation 5

```{r, comment=NA}
df$new_scores <- rnorm(n=nrow(df), mean=100, sd=5)
df
```

# Problem 7

What is the R statement used to install a package?

```{r,comment=NA}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
```

Load a package? 

```{r,comment=NA}
library(ggplot2)
```

Should the package be installed every time it is loaded?

No, the packaged should only be installed once and can be loaded multiple times.

# Problem 8

Write an R function normalize_0_1 that normalizes a numeric vector or matrix between 0 and 1 using the following formula:

$$\frac{x-min(x)}{max(x)-min(x)}$$

```{r, comment=NA}
normalize_0_1 <- function(input){
	min <- min(input)
	max <- max(input)
	input <- (input-min)/(max-min)
	
	return (input)
}
```

Use normalize_0_1 to normalize a vector v with 10 uniformly distributed random numbers between 100 and 200.

```{r,comment=NA}
v <- runif(10, min=100, max=200)

normalize_0_1(v)
```