---
title: "Week 5 Quiz - ggplot2"
output: html_notebook
---

1. Using set.seed(1187) generate a 1000 row sample (without replacement) of the ggplot2::diamonds dataset, and create a plot showing price by carat, using color to show how cut influences price, also using size to show how depth influences price. (scatterplot)

```{r}
library(ggplot2)
library(dplyr)
diamonds %>% glimpse

set.seed(1187)
samples <- sample_n(diamonds,
       size= 1000, 
       replace = FALSE)

ggplot(samples, aes(x=carat, y=price)) + 
  geom_point(aes(color=as.factor(cut), size=depth))
```

2. Use the dataset BitCoin_Data.csv located in Files/Class Notes/Week 5 to generate a plot showing log(mean_price) by Date using a red line graph, where the total volume for each point is indicated by the size of the point.

```{r}
library(tidyverse)
library(lubridate)
bt <- read_csv("C:\\Users\\lye\\Downloads\\BitCoin_data.csv")
bt$Date <- mdy(bt$Date)

bt_summary <- bt %>%
  group_by(Date) %>%
  summarise(log_mean_price = log10(`24h Average`), volume = sum(`Total Volume`))

ggplot(bt_summary, aes(x=Date, y=log_mean_price)) + 
  geom_point(aes(size=volume)) + #static
    geom_line(color="red") #dynamic

```

3. 