---
title: "Quiz Week 3"
output:
  html_document: default
  html_notebook: default
---

Use the nycflights13::flights to answer the following questions. Use diplyr pipes in all your answers.

1. Create two data frames, one consisting of 1000 observations, the other consisting of the 0.1% of the flights dataset.

```{r, message=FALSE}
library(nycflights13)
library(dplyr)

f1000 <- sample_n(nycflights13::flights, 1000)
f0001 <- sample_frac(nycflights13::flights, 0.001)
```

2. Create a data frame that removes duplicate rows based on departure day and carriers only.

```{r, message=FALSE}
new_df <- nycflights13::flights %>%
  distinct(day, carrier, .keep_all=TRUE)
```

3. Create a data frame from df1 consisting of only carrier and tailnum, create a data frame df2 consisting of all of the original variables in flights except carrier and tailnum.

```{r, message=FALSE}
df1 <- select(nycflights13::flights, carrier, tailnum)
df2 <- select(nycflights13::flights, -matches("carrier", "tailnum"))

df1 <- nycflights13::flights %>%
  select(carrier, tailnum)
df2 <- nycflights13::flights %>%
  select(-carrier, -tailnum)
```

Select function can also be used to reorder fields in the dataframe.

4. Create a data frame df consisting of only the fields related to time.

```{r, message=FALSE}
df <- nycflights13::flights %>%
  select(matches("time"))
```

Can use regular expressions as well.

5. Create a dataframe with flight number as the first column followed by the other columns in flights.

```{r, message=FALSE}
df <- nycflights13::flights %>%
  select(flight, everything())
```

6. Create a data frame where the column tailnum is changed to tail_number.

```{r, message=FALSE}
df <- nycflights13::flights %>%
  rename(tail_number = tailnum)
```

rename(new_Name=old_Name)

7. Create a data frame consisting of only flights originating from jfk airport and carrier is united airlines.

```{r, message=FALSE}
df <- nycflights13::flights %>%
  filter(origin == "JFK" & carrier == "UA")
```

&& is bitwise and individualized element-wise comparison.
Use commas instead of &

8. What is the median and mean arrival delay for all flights? 

What is the median and mean arrival delay originating out of jfk by carrier?

```{r, message=FALSE}
nycflights13::flights %>% 
  summarise(mean_arr_delay = mean(arr_delay, na.rm=TRUE), median_arr_delay = median(arr_delay, na.rm=TRUE))
```

```{r, message=FALSE}
nycflights13::flights %>%
  filter(origin == "JFK") %>%
  group_by(carrier) %>%
  summarise(mean_arr_delay = mean(arr_delay, na.rm=TRUE), median_arr_delay = median(arr_delay, na.rm=TRUE))
```

na.rm=TRUE means to ignore the NA values

9. What are the three days in a month with the lowest maximum departure delay, and what is it?

```{r, message=FALSE}
nycflights13::flights %>%
  group_by(day) %>%
  summarise(maximum_departure = max(dep_delay, na.rm=TRUE)) %>%
  arrange(maximum_departure) %>%
  head(3)
```

top_n(-3) would work as well

10. Explain the difference in results between these two dplyr statements:

```{r, message=FALSE}
flights %>%
  mutate(mean_dep_delay=mean(dep_delay, na.rm=TRUE))
```

```{r, message=FALSE}
flights %>%
  summarise(mean_dep_delay = mean(dep_delay, na.rm=TRUE))
```

The first statement adds another column to the data frame with the mean of departure delay, whereas the second statement creates a separate table that summarizes the data frame, and the only column is the mean of the departure delay.

11. 

```{r, message=FALSE}

```

